---
title: "Stops Codebook"
format: pdf
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))
```

```{r}
stops <- my_read_csv(here("merge-stops-shifts", "input", "stops.csv"))
summary <- GetSummary(stops)
```

## Stops

* **stop_id**: Identifier for each stop.
* **time**: Time of the stop in YYYY-MM-DD HH:MM:SS format.
* **date**: Date of the stop in YYYY-MM-DD format.
* **district**: Police district where the stop took place.
* **po_first**: Was the focal officer the first to respond to the scene?
* **stop_type**: What was the type of the stop?
* **contact_type**: Collapsed version of **stop_type** (less categories).
* **civ.race**: Race of the civilian.
* **civ.gender**: Sex of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **lat**: Latitude of the stop.
* **lon**: Longitude of the stop.
* **officer_id**: Unique identifier for the officer.
* **month**: Month of the stop in YYYY-MM-DD format. The day is always the first day of the month.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* **hour**: Hour of the day when the stop took place rounded to the nearest hour in military time.
* The unit of analysis is a unique officer involved in a stop. Each row can be uniquely identified by **officer_id** and **stop_id**.
* Each stop involves only one civilian, but they can involve multiple officers. It is possible multiple stops are all a part of one larger incident involving multiple civilians. This can be investigated by examining stops that took place in the same location at the same time involving the same officers.
* **Number of rows**: `r nrow(stops)`
* **Number of unique stops**: `r length(unique(stops$stop_id))`

### Stop Type vs. Contact Type

> "Stops for 'dispersal' and 'gang and narcotics-related loitering' are coded as loitering stops; those that are 'gang / narcotics related' are coded as drug stops; 'investigatory stops' and stops of 'suspicious persons' are coded as suspicious behavior; and stops under the 'Repeat Offender Geographic Urban Enforcement Strategy (ROGUES)' program are combined with the 'other' category." page 8 of Appendix Section S1.5 in Ba et al. 2021.

\newpage

```{r, results = "asis"}
summary[["numeric"]] %>% kable(format = "latex")
summary[["other"]] %>% kable(format = "latex")
summary[["logical"]] %>% kable(format = "latex")
summary[["factor"]] %>% kable(format = "latex", longtable = T)
```
