---
title: "Looking at the complaint civilian witness data"
output: html_document
---

```{r, message = F, warning = F}
library(here)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(lubridate)
```

**Complaints Civilian Witnesses**

* Each row captures a specific civilian witness - complaint pair.
* The same complaint can have multiple civilian witnesses.
* Witnesses don't have any ID associated with them.
* Not every complaint (in fact it seems very few complaints) actually have a civilian witness. It's unclear to me the process by which a civilian witness would get associated with a complaint.
* Not every civilian witness complaint has a match in the complaint accused (total) data set. I.e. some civilian witness complaints don't have a corresponding officer associated with the complaint.

```{r}
complaintsCivWitness18 <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-civilian-witnesses_2000-2018_2018-03.csv"),
             col_types = cols(complaint_date = col_date(format = "%Y-%m-%d"),
                              cr_id = "c")) %>%
    mutate(birth_year_date = ymd(birth_year, truncated = 3),
           age = floor(time_length(difftime(complaint_date, birth_year_date),
                                   "years")))

complaintsAccusedTotal <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-accused.csv"),
             col_types = cols(recc_finding = "c", recc_outcome = "c",
                              cr_id = "c")) %>%
    mutate(UID = as.character(UID)) %>%
    mutate(final_outcome = tolower(final_outcome),
           recc_outcome = tolower(recc_outcome))
```

Number of unique civilian witness complaints: `r nrow(complaintsCivWitness18)`  
Number of total civilian witness complaints: `r length(unique(complaintsCivWitness18$cr_id))`

```{r}
match <- left_join(complaintsCivWitness18, complaintsAccusedTotal, by = "cr_id")
```

How many civilian complaints don't match with the complaint accused (cop level) data?: `r table(is.na(match$UID))[["TRUE"]]`  
Percentage of total complaint incidents which don't match: `r table(is.na(match$UID))[["TRUE"]] / nrow(match)`

```{r}
ggplot(complaintsCivWitness18, aes(x = fct_infreq(race))) +
    geom_bar(aes(y = ..count.. / sum(..count..))) +
    labs(x = "Race", y = "Percentage",
         title = "Percentage of civilian complaints by race") +
    coord_flip() +
    theme_bw()
```

```{r}
ggplot(complaintsCivWitness18, aes(x = gender)) +
    geom_bar(aes(y = ..count.. / sum(..count..))) +
    labs(y = "Percentage") +
    theme_bw()
```

```{r}
ageDf <- complaintsCivWitness18 %>% filter(age > 0, age < 200)
ggplot(ageDf, aes(x = age)) + geom_histogram(bins = 40) + theme_bw()
```

```{r}
complaintDates <-
    complaintsCivWitness18 %>%
    distinct(cr_id, .keep_all = T) %>%
    mutate(year_complaint = year(complaint_date),
           month_complaint = month(complaint_date),
           day_complaint = day(complaint_date)) %>%
    select(year_complaint, month_complaint, day_complaint) %>%
    pivot_longer(everything(), "date_type") %>%
    count(date_type, value)

makeComplaintDatePlot <- function(df, date_t, date_label) {
    complaintDates %>% {
        filter(., date_type == date_t) %>%
        ggplot(aes(x = value, y = n)) +
        geom_line(stat = "identity") +
        labs(x = date_label,
             title = paste0("Complaints By ",
                            date_label,
                            " (Civilian Witnesses)")) +
        scale_x_continuous(breaks = seq(min(.$value), max(.$value), 2)) +
        theme_bw()}
}

makeComplaintDatePlot(complaintDates, "year_complaint", "Year")
makeComplaintDatePlot(complaintDates, "month_complaint", "Month")
makeComplaintDatePlot(complaintDates, "day_complaint", "Day")
```

