---
title: "Looking at the complaint accused data"
output: html_document
---

```{r, message = F, warning = F}
library(here)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
```

Complaints Accused: Each row captures a specific complaint - police officer pair. A complaint can be associated with multiple police officers.

Based on the following analyses, I plan on using the total file **complaints-accused.csv** which includes records from the three other component data sets (167-1999, 2000-2016, 2000-2018). The total data set, the 1967-1999 data set, and the 2000-2016 all pretty much have the same columns. The 2000-2018 data set has many more columns, but I'm still too unsure of how it relates to the 2000-2016 data set to feel confident using it.  

Based on conversations I had with the Invisible Institute, I plan on using the link_UID for uniquely identifying police officers. In this data set, they have a one-to-one relationship.

```{r}
# Have to use mutate to convert UID to character otherwise a trailing 0 appears.
complaintsAccusedTotal <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-accused.csv"),
             col_types = cols(recc_finding = "c", recc_outcome = "c",
                              cr_id = "c")) %>%
    mutate(UID = as.character(UID),
           link_UID = as.character(link_UID),
           final_outcome = tolower(final_outcome),
           recc_outcome = tolower(recc_outcome))

complaintsAccused6799 <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-accused_1967-1999_2016-12.csv"),
             col_types = cols(rank_no = "c", cr_id = "c")) %>%
    mutate(UID = as.character(UID), link_UID = as.character(link_UID))

complaintsAccused0016 <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-accused_2000-2016_2016-11.csv"),
             col_types = cols(cr_id = "c")) %>%
    mutate(UID = as.character(UID), link_UID = as.character(link_UID))

complaintsAccused0018 <-
    read_csv(here("unified_data", "full-output", "complaints",
                  "complaints-accused_2000-2018_2018-03.csv"),
             col_types = cols(cr_id = "c")) %>%
    mutate(UID = as.character(UID), link_UID = as.character(link_UID))
```

Percentage of UID not missing in the total data set: `r table(complaintsAccusedTotal$link_UID == -999)["FALSE"] / nrow(complaintsAccusedTotal)`  
Percentage of UID not missing from 1967-1999: `r table(complaintsAccused6799$link_UID == -999)["FALSE"] / nrow(complaintsAccused6799)`  
Percentage of UID not missing from 2000-2016: `r table(complaintsAccused0016$link_UID == -999)["FALSE"] / nrow(complaintsAccused0016)`  
Percentage of UID not missing from 2000-2018: `r table(complaintsAccused0018$link_UID == -999)["FALSE"] / nrow(complaintsAccused0018)`

```{r}
# Complaints from 1967 - 1999 not included in total because officer ID is missing
antij6799 <- anti_join(complaintsAccused6799, complaintsAccusedTotal, by = "cr_id")
table(antij6799$link_UID == -999)

# No complaints from 2000 - 2016 are missing from the total
antij0016 <- anti_join(complaintsAccused0016, complaintsAccusedTotal, by = "cr_id")

# Complaints from 2000 - 2018 not included in total because officer ID is missing
antij0018 <- anti_join(complaintsAccused0018, complaintsAccusedTotal, by = "cr_id")
table(antij0018$link_UID == -999)

complaintsAccused6799Fltr <- complaintsAccused6799 %>% filter(link_UID != -999)
complaintsAccused0018Fltr <- complaintsAccused0018 %>% filter(link_UID != -999)
```

```{r}
# Cops and complaints match which is good
cops6799 <- left_join(complaintsAccused6799Fltr,
                      select(complaintsAccusedTotal, cr_id, link_UID),
                      by = c("cr_id", "link_UID"))

cops0016 <- left_join(complaintsAccused0016,
                      select(complaintsAccusedTotal, cr_id, link_UID),
                      by = c("cr_id", "link_UID"))

cops0018 <- left_join(complaintsAccused0018Fltr,
                      select(complaintsAccusedTotal, cr_id, link_UID),
                      by = c("cr_id", "link_UID"))
```

Do all cop/complaint pairs match in the 1967-1999 data set with the total dataset?: `r nrow(cops6799) == nrow(complaintsAccused6799Fltr)`  
Do all cop/complaint pairs match in the 2000-2016 data set with the total dataset?: `r nrow(cops0016) == nrow(complaintsAccused0016)`  
Do all cop/complaint pairs match in the 2000-2018 data set with the total dataset?: `r nrow(cops0018) == nrow(complaintsAccused0018Fltr)`

* There are complaints from the 2016 data set not in the 2018 data set. The only reason I can think of is that these complaints don't have a newer 2018 entry which has more columns.
* There are some complaints in the 2018 data set not in the 2016 data set which makes theoretical sense because those complaints could have occurred after 2016.
* There are also some complaints in the 2016 and 2018 records which have different link_UIDs. Some of the cr_id + link_UID pairs are used from 2016 while others are used from 2018 in the total data set.
    * If there's an entry for link_UID.x and link_UID.y, that means the 2016 and 2018 sources disagree. If there's only one entry (link_UID.x or link_UID.y) that means that particular cop appeared in one of the data sets but not the other. The data set with a missing value had every cop for that particular complaint show up in the other data set. E.g. cr_id 1000674 has a missing value for link_UID.x (2016) which means every cop that showed up on that complaint in 2016 also showed up in 2018, but there was one cop in 2018 who doesn't show up on the 2016 version of that complaint.

```{r}
# Some 2016 complaints aren't found in 2018 
aj0016_0018 <-
    anti_join(complaintsAccused0016, complaintsAccused0018Fltr, by = "cr_id")

# Some 2018 complaints aren't found in 2016
aj0018_0016 <-
    anti_join(complaintsAccused0018Fltr, complaintsAccused0016, by = "cr_id")

# Match on cr_id, find those UIDs which don't match (2016 to 2018) 
copDiff0016_0018 <-
    inner_join(select(complaintsAccused0016, cr_id, link_UID),
               select(complaintsAccused0018Fltr, cr_id, link_UID),
               by = "cr_id") %>%
    group_by(cr_id, link_UID.x) %>%
    filter(all(link_UID.x != link_UID.y)) %>%
    distinct(cr_id, link_UID.x)

# Match on cr_id, find those UIDs which don't match (2018 to 2016)
copDiff0018_0016 <-
    inner_join(select(complaintsAccused0016, cr_id, link_UID),
               select(complaintsAccused0018Fltr, cr_id, link_UID),
               by = "cr_id") %>%
    group_by(cr_id, link_UID.y) %>%
    filter(all(link_UID.y != link_UID.x)) %>%
    distinct(cr_id, link_UID.y)

copDiff <- full_join(copDiff0016_0018, copDiff0018_0016, by = c("cr_id"))

# Complaints with differing cops are pulled from both 2016 and 2018 sources
copDiffTotal <- filter(complaintsAccusedTotal, cr_id %in% unique(copDiff$cr_id))
```

Confusingly there can be multiple rows for the same complaint + police officer combination. The 2018 data set in particular has an enormous amount of these instances suggesting there (at least for the 2018 data set) there is another unit of analysis being captured. The total data set and the 2016 data set are thankfully have no duplicate complaint + police officer rows.

```{r}
dupeCrUidT <- complaintsAccusedTotal %>% count(cr_id, link_UID)
dupeCrUid6799 <- complaintsAccused6799Fltr %>% count(cr_id, link_UID)
dupeCrUid0016 <- complaintsAccused0016 %>% count(cr_id, link_UID)
dupeCrUid0018 <- complaintsAccused0018Fltr %>% count(cr_id, link_UID)
```

**complaints-accused.csv**

* **final_finding**
    * DIS - ?
    * EX - Exonerated
    * NAF - No Affidavit
    * NC - ?
    * NS - Not sustained
    * SU - Sustained
    * UN - Unfounded

```{r}
ggplot(complaintsAccusedTotal, aes(x = fct_infreq(final_finding))) +
    geom_bar(aes(y = ..count.. / sum(..count..))) +
    labs(x = "Final Finding Categories", y = "Percentage") +
    theme_bw()
```

```{r}
findings <-
    complaintsAccusedTotal %>%
    filter(!is.na(recc_finding)) %>%
    select(final_finding, recc_finding) %>%
    pivot_longer(cols = everything(),
                 names_to = "finding_type",
                 values_to = "findings") %>%
    count(finding_type, findings) %>%
    group_by(finding_type) %>%
    mutate(percentage = n / sum(n)) %>%
    ungroup()

ggplot(findings, aes(x = fct_reorder(findings, percentage), y = percentage)) +
    geom_bar(stat = "identity") +
    facet_wrap(~finding_type) +
    labs("Finding Categories", "Recommended Findings vs. Final Findings") +
    theme_bw()
```

```{r}
copComplaints <- complaintsAccusedTotal %>% count(link_UID)
ggplot(copComplaints, aes(x = n)) + geom_histogram(bins = 35) + theme_bw()
summary(copComplaints$n)
```

Other interesting things of note:

* I could dig into the complaint categories. There are nearly 200 complaint categories I think so it might be interesting to look at what are the most common types of complaints.
* Outcomes could be interesting. Once the investigation comes to a finding, an appropriate punishment has to be decided (or no punishment as per the result of the investigation). The main punishment categories include:
    * Suspension for a certain number of days (ranges from 1 to 180)
    * Administrative Termination
    * Resigned
    * Reprimand
    * Violation Noted
    * Separation
    * Reinstated by police board
* Do certain types of complaints never get punished? Or some complaints always get punished? What's the relationship look like between complaint type and punishment outcome?
* What does the punishment history look like for police in general?