---
title: "Cleaning stops, arrests, and uses of force"
output: html_document
---

```{r}
library(here)
library(dplyr)
library(readr)
```

## Officer stops

* **stop_id**: Uniquely identifier for each stop.
* **time**: Time of stop.
* **date**: Date of stop.
* **district**: Police district where the stop took place.
* **po_first**: Was the focal officer the first to respond to the scene?
* **stop_type**: What was the type of the stop?
* **contact_type**: Collapsed version of **stop_type** (less categories).
* **civ.race**: Race of the civilian.
* **civ.gender**: Gender of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **lat**: Latitude of the stop.
* **lon**: Longitude of the stop.
* **officer_id**: Unique identifier for the officer.
* **month**: Month of stop.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* **hour**: Hour of the day when the stop took place.
* The unit of analysis is a unique officer involved in a stop. Each row can be uniquely identified by **officer_id** and **stop_id**.
* Each stop involves only one civilian, but they can involve multiple officers. It is possible multiple stops are all a part of one larger incident involving multiple civilians. This can be investigated by examining stops that took place in the same location at the same time involving the same officers.

```{r}
stops <- read_csv(here("0-clean-data", "input", "bocar-ba_data", "stops.csv"))
```

