---
title: "Clean Officer Shift Assignments"
output: pdf_document
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))

read_path <- paste("clean-and-process", "output", sep = "/")
write_path <- paste("clean-and-process", "output", sep = "/")
`%.%` <- paste0
```

```{r}
officers <- my_read_csv(here(read_path, "cleaned-officers.csv"))
assignments <- my_read_csv(here(read_path, "cleaned-assignments.csv"))
setDT(officers)
setDT(assignments)

# merge officer characteristics
officer_assignments <- assignments[officers[,.(officer_id,
                                              birth_year,
                                              officer_race,
                                              officer_gender,
                                              spanish)],
                                  on = 'officer_id']

# drop officers for whom no assignments were found
officer_assignments <- officer_assignments[!is.na(date),]

# Find missing shift times
missing_time <- officer_assignments %>% filter(is.na(start_time)) %>% nrow()

# drop all but POLICE OFFICER
rank_fltr <- officer_assignments[rank == 'POLICE OFFICER',]

# drop all officers except black, Hispanic, and white
race_fltr <- officer_assignments[
    officer_race %in% ('officer_' %.% c('black', 'white', 'hisp')),]

# final data set
officer_assignments_ba <-
    officer_assignments %>%
    filter(officer_race %in% c(paste0("officer_", c("white", "black", "hisp"))),
           rank == "POLICE OFFICER")

# Find missing shift times
missing_time_ba <- officer_assignments_ba %>% filter(is.na(start_time)) %>% nrow()

# Find officers which had no shifts and any shifts which had no officers
anti_join_assignments <-
    anti_join(assignments, officers, by = "officer_id") %>%
    nrow()

anti_join_officers <-
    anti_join(officers, assignments, by = "officer_id") %>%
    nrow()

write_csv(officer_assignments, here(write_path, "officers_assignments_risi.csv"))
write_csv(officer_assignments_ba, here(write_path, "officers_assignments_ba.csv"))
```

## Merging Officers To Their Shift Assignments - Risi

* Number of officer shift assignments: `r nrow(officer_assignments)`
* Number of unique officers who had at least one shift assignment: `r length(unique(officer_assignments$officer_id))`
* Number of non-matching assignments: `r anti_join_assignments`
    * Percentage of assignments which don't match: `r anti_join_assignments / nrow(assignments) * 100`%
* Number of non-matching officers: `r anti_join_officers`
    * Percentage of officers who don't match: `r anti_join_officers / nrow(officers) * 100`%
* Of the `r nrow(officer_assignments)` shift assignments, `r missing_time` are missing their start time and end time.
    * This represents `r 100 * missing_time / nrow(officer_assignments_ba)`% of all shift assignments.

```{r, results = "asis"}
compare <-
    officers %>%
    mutate(has_a_shift = as.character(
        if_else(officer_id %in% officer_assignments$officer_id,
                1,
                0)))

 ggplot(compare, aes(x = birth_year)) +
    geom_histogram(aes(fill = has_a_shift)) +
    facet_wrap(~has_a_shift) +
    theme_bw()
 
compare1 <-
    GetSummaryCol(filter(compare, has_a_shift == 1), "birth_year") %>%
    mutate(shift = 1)

compare0 <-
    GetSummaryCol(filter(compare, has_a_shift == 0), "birth_year") %>%
    mutate(shift = 0)

kable(rbind(compare1, compare0), format = "latex")
```

It's unclear why so many officers don't have a shift assignment. Most of the non-matches are likely due to the fact that this is a roster of every person who ever served as an officer so it includes retired officers or officers who quit (as of when these shift assignments start). Additionally some of the officers listed in the roster likely do not get assigned to patrol shifts due to the nature of their role (e.g. they are in a leadership role).  

More worrisome might be officers who don't get shift assignments for some other, unknown reason which might hurt the validity of the study (e.g. they're being punished or rewarded).

Officers who do have at least one shift assignment are noticeably younger lending some support to the idea that officers who don't have a shift assignment have retired.

## Filtering - Ba et al. 2021

### Race

> "We restrict analysis to patrol assignments in which Black, Hispanic, or White officers serve. Asian/Pacific Islander and Native American/Alaskan Native officers are not examined due to small sample sizes." page 7 of Appendix Section S1.4 in Ba et al. 2021.

* Number of officer shift assignments: `r nrow(race_fltr)`
    * Shifts dropped: `r nrow(officer_assignments) - nrow(race_fltr)`
    * Percentage reduction: `r (nrow(officer_assignments) - nrow(race_fltr)) / nrow(officer_assignments) * 100`%
* Number of unique officers: `r length(unique(race_fltr$officer_id))`
    * Number Of Officers dropped: `r length(unique(officer_assignments$officer_id)) - length(unique(race_fltr$officer_id))`
    * Percentage reduction: `r (length(unique(officer_assignments$officer_id)) - length(unique(race_fltr$officer_id))) / length(unique(officer_assignments$officer_id)) * 100`%
    
### Rank

> "Finally, we drop officers at ranks other than 'police officer.' This step eliminates police sergeants, who serve in 8% of beat assignments but make very few stops and arrests, as well as legal officers, helicopter pilots, explosives technicians, and canine handlers." page 7 of Appendix Section S1.4 in Ba et al. 2021.

* Number of officer shift assignments: `r nrow(rank_fltr)`
    * Shifts dropped: `r nrow(officer_assignments) - nrow(rank_fltr)`
    * Percentage reduction: `r (nrow(officer_assignments) - nrow(rank_fltr)) / nrow(officer_assignments) * 100`%
* Number of unique officers: `r length(unique(rank_fltr$officer_id))`
    * Officers dropped: `r length(unique(officer_assignments$officer_id)) - length(unique(rank_fltr$officer_id))`
    * Percentage reduction: `r (length(unique(officer_assignments$officer_id)) - length(unique(rank_fltr$officer_id))) / length(unique(officer_assignments$officer_id)) * 100`%
    
### Totals

Note the numbers above won't completely add up to the numbers below because of intersections.

* **Number of officer shift assignments**: `r nrow(officer_assignments_ba)`
    * Shifts dropped: `r nrow(officer_assignments) - nrow(officer_assignments_ba)`
    * Percentage reduction: `r (nrow(officer_assignments) - nrow(officer_assignments_ba)) / nrow(officer_assignments) * 100`%
* **Number of unique officers with at least one shift assignment**: `r length(unique(officer_assignments_ba$officer_id))`
    * Officers dropped: `r length(unique(officer_assignments$officer_id)) - length(unique(officer_assignments_ba$officer_id))`
    * Percentage reduction: `r (length(unique(officer_assignments$officer_id)) - length(unique(officer_assignments_ba$officer_id))) / length(unique(officer_assignments$officer_id)) * 100`%
* Of the `r nrow(officer_assignments_ba)` shift assignments, `r missing_time_ba` are missing their start time and end time.
    * This represents `r 100 * missing_time_ba / nrow(officer_assignments_ba)`% of all shift assignments.

## Visualizing Results

### What are the demographic breakdowns of active officers?

* Racially, about 50% of all active patrol officers are White. 25% are Black. 20% are Hispanic.
* 75% of officers are male.
* 85% do not speak Spanish.
* When looking at race and gender, Native Americans, White Americans, and Hispanic Americans have roughly an 80/20 split when it comes to male/female representation. Black Americans actually have a 65/35 split and Asian Americans & Pacific Islanders have a nearly 90/10 split. Interestingly, white women are not the majority among women but a plurality.
* Not surprisingly, Hispanic Americans are the most common racial group among those who speak Spanish. Roughly 46.5% of all Hispanic officer speak Spanish while most other racial groups have around 5-10% of their groups being able to speak Spanish.

```{r}
officer_level <-
    officer_assignments %>%
    count(officer_id, officer_race, officer_gender, spanish)

officer_level %>%
    select(-n, -officer_id) %>%
    filter(!is.na(officer_race)) %>%
    mutate(officer_race = str_replace(officer_race, "officer_", "")) %>%
    ggpairs(upper = list(discrete = "colbar"), lower = list(discrete = "colbar"))
```

How many shift assignments do officers typically get assigned to?

```{r}
ggplot(officer_level, aes(x = n)) +
    geom_histogram(bins = 40) +
    theme_bw() +
    labs(x = "Number of Shift Assignments", y = "Number of Officers")

GetSummaryCol(officer_level, "n")
```



