---
title: "Description of the Data"
output: pdf_document
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, fig.height = 8, fig.width = 12,
               out.width='\\textwidth')

library(here)
source(here("functions.R"))

library(kableExtra)

final_vars <-
    read_csv(here("clean-and-process", "src", "old_stuff", "final_data.csv")) %>%
    mutate(across(c("unit", "year", "unit"), ~as.factor(.x))) %>%
    mutate(log_total_officers = log(nr_officer),
           log_black_officers = log(nr_officer_black),
           log_black = log(black),
           property_cr_capita = property_cr * 10000 / total_pop,
           violent_cr_capita = violent_cr * 10000 /  total_pop,
           mean_years_exp = mean_days_exp / 365)
```

## Coefficient of variation

```{r}
coev_dep <-
    final_vars %>%
    group_by(unit) %>%
    summarise(mean = mean(black_stops),
              sd = sd(black_stops),
              coev = sd / mean,
              median = median(black_stops),
              mad = mad(black_stops),
              mad_coev = mad / median)

coev_indep <-
    final_vars %>%
    group_by(unit) %>%
    summarise(mean = mean(black_ratio),
              sd = sd(black_ratio),
              coev = sd / mean,
              median = median(black_ratio),
              mad = mad(black_ratio),
              mad_coev = mad / median)
```

## Dependent Variable

### Within Each Unit

```{r}
ggplot(final_vars, aes(x = date, y = black_stops)) +
    facet_wrap(~unit, scales = "free_y") +
    geom_point() +
    theme_bw() +
    labs(x = "Date", y = "Number of Stops Of Black Civilians",
         title = "Number of Stops Within Each Unit Over Time") +
    theme(text = element_text(size = 16))
```

### Between Each Unit

```{r}
ggplot(final_vars, aes(x = date, y = black_stops)) +
    geom_point(aes(color = unit)) +
    geom_line(aes(color = unit, group = unit)) +
    theme_bw() +
    labs(x = "Date", y = "Number of Stops Of Black Civilians",
         title = "Number of Stops Between Each Unit Over Time") +
    theme(text = element_text(size = 16))

ggplot(final_vars, aes(x = date, y = black_stops)) +
    geom_point() +
    geom_line(aes(group = unit)) +
    facet_wrap(~unit) +
    theme_bw() +
    labs(x = "Date", y = "Number of Stops Of Black Civilians",
         title = "Number of Stops Between Each Unit Over Time")
```

## Main Independent Variable - Black Racial Congruence

### Overall distribution of Black Racial Congruence

* Most police units in most months have a black congruence below 1 meaning the share of the population that is Black is greater than the share of the police force which is Black.

```{r}
ggplot(final_vars, aes(x = black_ratio)) +
    geom_histogram(bins = 15) +
    theme_bw() +
    labs(x = "Black Racial Congruence", y = "Count",
         title = "Distribution of Racial Congruence Across Units And Time")
```

### Within Each Unit

```{r}
ggplot(final_vars, aes(x = date, y = black_ratio)) +
    facet_wrap(~unit, scales = "free_y") +
    geom_point() +
    theme_bw() +
    labs(x = "Date", y = "Black Racial Congruence",
         title = "Black Racial Congruence Within Each Unit Over Time") +
    theme(text = element_text(size = 16))
```

### Between Each Unit

```{r}
ggplot(final_vars, aes(x = date, y = black_ratio)) +
    facet_wrap(~unit) +
    geom_point() +
    theme_bw() +
    labs(x = "Date", y = "Black Racial Congruence",
         title = "Black Racial Congruence Between Each Unit Over Time") +
    theme(text = element_text(size = 16))

ggplot(final_vars, aes(x = date, y = black_ratio)) +
    geom_point(aes(color = unit, group = unit)) +
    geom_line(aes(color = unit, group = unit)) +
    theme_bw() +
    labs(x = "Date", y = "Black Racial Congruence",
         title = "Black Racial Congruence Between Each Unit Over Time")
```

### What's the connection between the number of black officers in a unit and racial congruence?

* Generally within units, as more black officers are added to the unit, racial congruence increases. For some very large units the relationship is less strong which makes sense. When the share of civilians in the population who are Black is relatively large **and/or** when the size of the police unit is large, many more Black officers would have to join the police unit vs. another unit which was smaller or where the share of the civilian population which is Black is smaller.
* This distills why I use Black Racial Congruence as a measure. Adding 1 Black officer to a unit should *theoretically* have different effects depending on the unit and the population being policed.

```{r}
ggplot(final_vars, aes(x = black_ratio, y = nr_officer_black)) +
    geom_point(aes(color = unit)) +
    geom_line(aes(color = unit, group = unit)) +
    theme_bw() +
    labs(x = "Black Racial Congruence", y = "Number of Black Officers")
```

### Table demonstrating variation in black racial congruence within each unit

* Each row represents the maximum or minimum Black Racial Congruence achieved by each police unit during the period of observation (2013 - 2015).
* Along with the maximum or minimum racial congruence, the total number of police officers, the total number of Black officers, the percentage of Black officers, and the percentage of the civilian population that is Black are also included. It is designed to help the reader (and me) get a feel for what increasing Black racial congruence looks like. Typically it means adding more Black officers to the unit. Bigger increases come when the rate at which Black officers are being added is greater than the rate at which the total size of force is growing (or if the force is shrinking). Occasionally Black Racial Congruence can increase even when the number of Black officers is decreasing, and this is because the rate of decrease for the number of Black officers is less than the rate of decrease for the unit as a whole.
* Increasing Black Racial Congruence means an increase in the share of the police force that is Black relative to the share of the civilian population which is Black.

```{r, results = "asis"}
variation <-
    final_vars %>%
    group_by(unit) %>%
    filter(black_ratio == max(black_ratio) | black_ratio == min(black_ratio)) %>%
    mutate(value = if_else(black_ratio == min(black_ratio), "Min", "Max")) %>%
    ungroup() %>%
    dplyr::select(unit, nr_officer, nr_officer_black, prcnt_officer_black,
                  prcnt_civ_black, black_ratio, value) %>%
    distinct() %>%
    arrange(unit, desc(value))

kable(variation,
      booktabs = T,
      col.names = c("Unit",
                    "Officers",
                    "Black Officers",
                    "Pr. Black Officers",
                    "Pr. Black Pop.",
                    "Racial Congruence",
                    "Max/Min Congruence")) %>%
    kable_styling(full_width = F) %>%
    row_spec(seq(2, 42, by = 2), hline_after = T)
```
