---
title: "Use of Force Codebook"
format: pdf
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))
```

```{r}
force <- my_read_csv(here("merge-force-shifts", "input", "force.csv"))
summary <- GetSummary(force)
```

## Uses of Force

* **date**: Date of the use of force in YYYY-MM-DD format.
* **time**: Time of the use of force in YYYY-MM-DD HH:MM:SS format.
* **district**: Police district where the use of force took place.
* **lat**: Latitude of the use of force.
* **lon**: Longitude of the force.
* **civ.race**: Race of the civilian.
* **civ.gender**: Sex of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **civ.injured**: Was the civilian injured?
* **force_id**: Unique identifier for each use of force incident.
* **officer_id**: Unique identifier for each officer.
* **month**: Month of the arrest in YYYY-MM-DD format. The day is always the first day of the month.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* **hour**: Hour of the day when the use of force took place rounded to the nearest hour in military time.
* The unit of analysis is a use of force incident. Only one police officer is listed for each use of force incident. Each row can be uniquely identified by **force_id**.
* **Number of uses of force**: `r nrow(force)`

\newpage

```{r, results = "asis"}
summary[["numeric"]] %>% kable(format = "latex")
summary[["other"]] %>% kable(format = "latex")
summary[["factor"]] %>% kable(format = "latex", longtable = T)
```
