---
title: "Regression Analysis - Revisions"
header-includes: \usepackage{booktabs} \usepackage{siunitx} \newcolumntype{d}{S[input-symbols = ()]}
output: pdf_document
---

```{r, echo = F}
library(knitr)
opts_chunk$set(message = F, echo = F, out.width='\\textwidth')
options("modelsummary_format_numeric_latex" = "plain")
```

```{r}
library(plm)
library(pglm)
library(lme4)
library(MASS)
library(lme4)
library(vtable)
library(fixest)
library(lmtest)
library(GGally)
library(sandwich)
library(parameters)
library(modelsummary)

library(here)
source(here("functions.R"))

final_vars <-
    read_csv(here("paper-analyses", "input", "final_data.csv")) %>%
    mutate(across(c("unit", "year", "unit"), ~as.factor(.x))) %>%
    mutate(log_total_officers = log(nr_officer),
           log_black_officers = log(nr_officer_black),
           white_stop_rate = round(white_stops / white * 10000),
           hispanic_stop_rate = round(hispanic_stops / black * 10000),
           log_black = log(black),
           property_cr_capita = property_cr * 10000 / total_pop,
           violent_cr_capita = violent_cr * 10000 /  total_pop,
           mean_years_exp = mean_days_exp / 365,
           mean_years_worked_unit = mean_months_worked_unit / 12,
           percentage_officer_black = prcnt_officer_black * 100,
           percentage_civ_black = prcnt_civ_black * 100)
```

## Corrrelation Matrix

The new way of measuring experience and racial congruence are no longer correlated.

```{r}
corr_vars <-
    final_vars %>%
    dplyr::select(black_stops, black_stop_rate, nr_officer, black, black_ratio,
                  mean_years_worked_unit, violent_cr_capita, property_cr_capita) %>%
    rename(`Black Stops` = black_stops,
           `Black Stopping Rate` = black_stop_rate,
           `Nr. of Officers` = nr_officer,
           `Black Pop.` = black,
           `Black Racial Congruence` = black_ratio,
           `Years Worked In Unit (Mean)` = mean_years_worked_unit,
           `Violent Crime Per 10,000` = violent_cr_capita,
           `Property Crime Per 10,000` = property_cr_capita)
 
ggcorr(corr_vars,
       low = "#F21A00",
       high = "#3B9AB2",
       label = T,
       label_round = 2,
       hjust = 0.75,
       size = 2)
```

## Summary Table

```{r, results = "asis"}
table_vars <-
    final_vars %>%
    dplyr::select(stops, black_stops, black_stop_rate, white_stop_rate,
                  hispanic_stop_rate, nr_officer, nr_officer_black,
                  percentage_officer_black, total_pop, black, percentage_civ_black, 
                  black_ratio, mean_years_worked_unit, violent_cr_capita,
                  property_cr_capita)

labels_df <-
    tibble(vars = colnames(table_vars),
           labels = c("Total Number of Stops",
                      "Stops of Black Civilians",
                      "Stops of Black Civilians Per 10,000",
                      "Stops of White Civilians Per 10,000",
                      "Stops of Hispanic Civilians Per 10,000",
                      "Total Number of Officers",
                      "Total Number of Black Officers",
                      "Percent of Officers - Black",
                      "Total Population",
                      "Black Population",
                      "Percent of Population - Black",
                      "Black Racial Congruence",
                      "Years Worked In Unit (Mean)",
                      "Violent Crime Per 10,000 individuals",
                      "Property Crime Per 10,000 individuals"))

st(table_vars,
   summ = list(c('notNA(x)', 'mean(x)',   'sd(x)', 'min(x)', 'pctile(x)[25]', 
                 'median(x)', 'pctile(x)[75]', 'max(x)')),
   summ.names = list(c("N", "Mean", "Std. Dev.", "Min", "25th", "Median",
                     "75th", "Max")),
   labels = labels_df,
   title = "Summary Statistics Across All Units Over All Years")
```

## Black Individuals are Underrepresented in Nearly Every Police Unit

```{r, fig.width = 14, fig.height = 8}
final_vars %>%
    filter(month == "6") %>%
    ggplot(aes(x = prcnt_civ_black, y = prcnt_officer_black)) +
    geom_point() +
    geom_abline() +
    facet_wrap(~year) +
    theme_bw() +
    labs(x = "Percentage of Population - Black",
         y = "Percentage of Police Unit - Black",
         title = "% of the population that is Black vs. % of the police force that is Black") +
    theme(text = element_text(size = 12))
```

## The Median Stopping Rate is Highest for Black Individuals In Half of All Units

```{r, fig.width = 12, fig.height = 7}
graph <-
    final_vars %>%
    dplyr::select(unit, year, month, white_stop_rate, black_stop_rate, hispanic_stop_rate,
                  black, white, hispanic) %>%
    group_by(unit) %>%
    summarise(white_stop_rate = median(white_stop_rate),
              black_stop_rate = median(black_stop_rate),
              hispanic_stop_rate = median(hispanic_stop_rate)) %>%
    ungroup() %>%
    pivot_longer(cols = matches("rate"), names_to = "Race", values_to = "Stops") %>%
    mutate(Race = case_when(Race == "white_stop_rate" ~ "White",
                            Race == "black_stop_rate" ~ "Black",
                            Race == "hispanic_stop_rate" ~ "Hisp.")) %>%
    group_by(unit) %>%
    ungroup() %>%
    group_by(unit) %>%
    mutate(max_match = Stops == max(Stops),
           max_match_black = if_else(max_match & Race == "Black", T, F),
           max_unit = any(max_match_black)) %>%
    ungroup()

sorting_var <-
    graph %>%
    filter(Race == "Black") %>%
    arrange(desc(max_match), desc(Stops))

graph <- graph %>% mutate(unit = factor(unit, levels = sorting_var$unit))

graph %>%
    ggplot(aes(x = Race, y = Stops)) +
    geom_bar(stat = "identity", aes(fill = max_unit)) +
    theme_bw() +
    labs(x = "Race", y = "Stopping Rate Per 10,000",
         title = "Median Stopping Rate By Race For Each Police Unit",
         fill = "Stopping Rate Highest For Black Individuals?") +
    facet_wrap(~unit, scales = "free_y") +
    theme(text = element_text(size = 12))
```

## Between vs. Within

```{r, results = "asis", warning = F}
final_vars  <-
    final_vars %>%
    group_by(unit) %>%
    mutate(between_black_ratio = mean(black_ratio),
           within_black_ratio = black_ratio - between_black_ratio) %>%
    ungroup()

between_within_count <-
    femlm(black_stops ~ between_black_ratio + within_black_ratio + year,
          family = "negbin",
          data = final_vars,
          cluster = "unit")

between_within_rate <-
    femlm(black_stops ~ between_black_ratio + within_black_ratio + year +
              offset(log(black)),
          family = "negbin",
          data = final_vars,
          cluster = "unit")

offset_row <- tibble(term = c("", "Offset - Black Pop."),
                     `Model 1` = c("Model A1", "No"),
                     `Model 2` = c("Model A2", "Yes"))
attr(offset_row, "position") <- c(1, 7)

modelsummary(list("Stops of Black Civilians" = between_within_count,
                  "Stops of Black Civilians" = between_within_rate),
             coef_omit = "(Intercept)|theta",
             coef_rename = c("between_black_ratio" = "Black Racial Congruence - Between Unit",
                            "within_black_ratio" = "Black Racial Congruence - Within Unit",
                            "black_stops" = "Stops of Black Civilians"),
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             stars = T,
             gof_omit = "R2$|R2 Adj|R2 Within",
             exponentiate = T,
             output = "latex",
             add_rows = offset_row,
             notes = c("Standard Errors in parentheses. Coefficients are incident rate ratios.",
                       "P-values are denoted by symbols: + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"))
```

## Plotting Racial Congruence Vs. Rate of Stops

```{r, fig.width = 10}
ggplot(final_vars, aes(y = black_stop_rate, x = black_ratio)) +
    geom_point(aes(color = unit)) +
    geom_smooth(method = MASS::glm.nb,
                formula = y ~ x,
                se = F,
                aes(group = unit, color = unit)) +
    geom_smooth(method = MASS::glm.nb,
                formula = y ~ x,
                se = F,
                aes(group = 1)) +
    theme_bw() +
    labs(x = "Black Racial Congruence",
         y = "Stops of Black Civilians Per 10,000") +
    theme(text = element_text(size = 12))
```

## Regression Models

```{r, results = "asis"}
fixest_cluster_nb_prcnt_basic <-
    femlm(black_stops ~ prcnt_officer_black + violent_cr_capita +
              property_cr_capita + log_total_officers + year +
              offset(log(black)) | unit,
          family = "negbin",
          data = final_vars)

fixest_cluster_nb_prcnt_full <-
    femlm(black_stops ~ prcnt_officer_black + mean_years_worked_unit +
              violent_cr_capita + property_cr_capita + log_total_officers +
              year + offset(log(black)) | unit,
          family = "negbin",
          data = final_vars)

fixest_cluster_nb_basic <-
    femlm(black_stops ~ black_ratio + violent_cr_capita + property_cr_capita +
              log_total_officers + year + offset(log(black)) | unit,
          family = "negbin",
          data = final_vars)

fixest_cluster_nb_full <-
    femlm(black_stops ~ black_ratio + mean_years_worked_unit +
              violent_cr_capita + property_cr_capita + log_total_officers +
              year + offset(log(black)) | unit,
          family = "negbin",
          data = final_vars)

offset_row <- tibble(term = c("", "Offset - Black Pop."),
                     `Model 1` = c("Model 1", "Yes"),
                     `Model 2` = c("Model 2", "Yes"),
                     `Model 3` = c("Model 3", "Yes"),
                     `Model 4` = c("Model 4", "Yes"))
attr(offset_row, "position") <- c(1, 10)
rename = c(black_ratio = "Black Racial Congruence",
           prcnt_officer_black = "Percentage of Officers Who Are Black",
           mean_years_worked_unit = "Years Worked In Unit (Mean)",
           violent_cr_capita = "Violent Crime Per 10,000",
           property_cr_capita = "Property Crime Per 10,000",
           log_total_officers = "Log of the Total Number of Officers")

modelsummary(list("Stops of Black Civilians" = fixest_cluster_nb_basic,
                  "Stops of Black Civilians" = fixest_cluster_nb_prcnt_basic,
                  "Stops of Black Civilians" = fixest_cluster_nb_full,
                  "Stops of Black Civilians" = fixest_cluster_nb_prcnt_full),
             coef_omit = "(Intercept)|theta",
             coef_map = rename,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             stars = T,
             gof_omit = "R2$|R2 Adj|R2 Within",
             exponentiate = T,
             output = "latex",
             add_rows = offset_row,
             notes = c("Standard Errors in parentheses. Coefficients are incident rate ratios.",
                       "P-values are denoted by symbols: + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001")) %>%
    kable_styling(full_width = F, latex_options = "scale_down")
```

## Investigating thresholds

```{r}
# final_vars <-
#     final_vars %>%
#     mutate(percent10 = prcnt_officer_black >= 0.1,
#            percent15 = prcnt_officer_black >= 0.15,
#            percent20 = prcnt_officer_black >= 0.2,
#            percent25 = prcnt_officer_black >= 0.25,
#            percent30 = prcnt_officer_black >= 0.3)
# 
# fixest_cluster_nb_t10 <-
#     femlm(black_stops ~ black_ratio + mean_years_worked_unit + percent10 +
#               violent_cr_capita + property_cr_capita + log_total_officers +
#               year + offset(log(black)) | unit,
#           family = "negbin",
#           data = final_vars)
# 
# fixest_cluster_nb_t15 <-
#     femlm(black_stops ~ black_ratio + mean_years_worked_unit + percent15 +
#               violent_cr_capita + property_cr_capita + log_total_officers +
#               year + offset(log(black)) | unit,
#           family = "negbin",
#           data = final_vars)
# 
# fixest_cluster_nb_t20 <-
#     femlm(black_stops ~ black_ratio + mean_years_worked_unit + percent20 +
#               violent_cr_capita + property_cr_capita + log_total_officers +
#               year + offset(log(black)) | unit,
#           family = "negbin",
#           data = final_vars)
# 
# fixest_cluster_nb_t25 <-
#     femlm(black_stops ~ black_ratio + mean_years_worked_unit + percent25 +
#               violent_cr_capita + property_cr_capita + log_total_officers +
#               year + offset(log(black)) | unit,
#           family = "negbin",
#           data = final_vars)
# 
# fixest_cluster_nb_t30 <-
#     femlm(black_stops ~ black_ratio + mean_years_worked_unit + percent30 +
#               violent_cr_capita + property_cr_capita + log_total_officers +
#               year + offset(log(black)) | unit,
#           family = "negbin",
#           data = final_vars)
# 
# offset_row <- tibble(term = c("", "Offset - Black Pop."),
#                      `Model 1` = c("Model 10%", "Yes"),
#                      `Model 2` = c("Model 15%", "Yes"),
#                      `Model 3` = c("Model 20%", "Yes"),
#                      `Model 4` = c("Model 25%", "Yes"),
#                      `Model 5` = c("Model 30%", "Yes"))
# attr(offset_row, "position") <- c(1, 12)
# 
# modelsummary(list("Stops of Black Civilians" = fixest_cluster_nb_t10,
#                   "Stops of Black Civilians" = fixest_cluster_nb_t15,
#                   "Stops of Black Civilians" = fixest_cluster_nb_t20,
#                   "Stops of Black Civilians" = fixest_cluster_nb_t25,
#                   "Stops of Black Civilians" = fixest_cluster_nb_t30),
#              coef_omit = "(Intercept)|theta",
#              coef_rename = c(black_ratio = "Black Racial Congruence",
#                              percent10 = "10% Threshold",
#                              percent15 = "15% Threshold",
#                              percent20 = "20% Threshold",
#                              percent25 = "25% Threshold",
#                              percent30 = "30% Threshold",
#                              mean_years_exp = "Mean Amount of Experience (Years)",
#                              violent_cr_capita = "Violent Crime Per 10,000",
#                              property_cr_capita = "Property Crime Per 10,000",
#                              log_total_officers = "Log of the Total Number of Officers"),
#              estimate = "{estimate} ({std.error}){stars}",
#              statistic = NULL,
#              stars = T,
#              gof_omit = "R2$|R2 Adj|R2 Within",
#              exponentiate = T,
#              output = "latex",
#              add_rows = offset_row,
#              notes = c("Standard Errors in parentheses. Coefficients are incident rate ratios.",
#                        "P-values are denoted by symbols: + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"))
```

## Predictions To Aid In Interpretation of Coefficients

* Standardized coefficients do not make much sense to use in aiding in the interpretation of panel regressions. Should one standardize using the **total** mean and standard deviation? Within time unit mean and standard deviation? Within police unit? Within police unit crossed by time unit? Furthermore such interpretations in the context of a fixed effects regression remain unclear.
* To aid in the interpretation of the regression models, I systematically vary the value of the Black Racial Congruence variable (e.g. using maximum and minimum observed values for each unit within each year) while holding all other variables constant at their mean. I then do the same for the Years Worked In Unit (Mean) as a comparison point.

```{r}
nr_officers <-
    final_vars %>%
    dplyr::select(unit, year, black_ratio, nr_officer_black) %>%
    group_by(unit, year) %>%
    mutate(rank = row_number(black_ratio)) %>%
    mutate(type = case_when(rank == min(rank) ~ "min",
                            rank == max(rank) ~ "max",
                            rank == 4 ~ "p25",
                            rank == 6 ~ "med",
                            rank == 9 ~ "p75",
                            T ~ "drop")) %>%
    ungroup() %>%
    filter(type != "drop") %>%
     dplyr::select(-black_ratio, -rank)

new_data_ratio <-
    final_vars %>%
    group_by(unit, year, black) %>%
    summarise(prcnt_officer_FEMALE = mean(prcnt_officer_FEMALE),
              mean_years_worked_unit = mean(mean_years_worked_unit),
              violent_cr_capita = mean(violent_cr_capita),
              property_cr_capita = mean(property_cr_capita),
              log_total_officers = mean(log_total_officers),
              black = mean(black),
              min = min(black_ratio),
              max = max(black_ratio),
              med = median(black_ratio),
              p25 = quantile(black_ratio)[["25%"]],
              p75  = quantile(black_ratio)[["75%"]]) %>%
    pivot_longer(c("min", "max", "med", "p25", "p75"),
                 names_to = "type",
                 values_to = "black_ratio") %>%
    ungroup() %>%
    full_join(nr_officers, by = c("unit", "year", "type"))

new_data_exp <-
    final_vars %>%
    group_by(unit, year, black) %>%
    summarise(prcnt_officer_FEMALE = mean(prcnt_officer_FEMALE),
              black_ratio = mean(black_ratio),
              violent_cr_capita = mean(violent_cr_capita),
              property_cr_capita = mean(property_cr_capita),
              log_total_officers = mean(log_total_officers),
              black = mean(black),
              min = min(mean_years_worked_unit),
              max = max(mean_years_worked_unit),
              med = median(mean_years_worked_unit),
              p25 = quantile(mean_years_worked_unit)[["25%"]],
              p75 = quantile(mean_years_worked_unit)[["75%"]]) %>%
    pivot_longer(c("min", "max", "med", "p25", "p75"),
                 names_to = "type",
                 values_to = "mean_years_worked_unit") %>%
    ungroup()

new_data_prop <-
    final_vars %>%
    group_by(unit, year, black) %>%
    summarise(prcnt_officer_FEMALE = mean(prcnt_officer_FEMALE),
              black_ratio = mean(black_ratio),
              violent_cr_capita = mean(violent_cr_capita),
              mean_years_worked_unit = mean(mean_years_worked_unit),
              log_total_officers = mean(log_total_officers),
              black = mean(black),
              min = min(property_cr_capita),
              max = max(property_cr_capita),
              med = median(property_cr_capita),
              p25 = quantile(property_cr_capita)[["25%"]],
              p75 = quantile(property_cr_capita)[["75%"]]) %>%
    pivot_longer(c("min", "max", "med", "p25", "p75"),
                 names_to = "type",
                 values_to = "property_cr_capita") %>%
    ungroup()
```

### The effect of moving from the least amount of racial congruence to the most for each year for each unit

As mentioned above, I hold all other values constant at their means (for each unit in each year). I then systematically vary the value of Black racial congruence using the minimum value (for that unit in that year), the 25th percentile, the median, the 75th percentile, and the maximum value.

```{r, fig.width = 14, fig.height = 9}
predict_results_ratio <-
    new_data_ratio %>%
    mutate(predicted_black_stops = predict(fixest_cluster_nb_full,
                                           newdata = new_data_ratio) * black,
           type = factor(type, levels = c("min", "p25", "med", "p75", "max")),
           variable = "Black Racial Congruence")

predict_results_exp <-
    new_data_exp %>%
    mutate(predicted_black_stops = predict(fixest_cluster_nb_full,
                                           newdata = new_data_exp) * black,
           type = factor(type, levels = c("min", "p25", "med", "p75", "max")),
           variable = "Years Worked In Unit (Mean)")

predict_results_prop <-
    new_data_prop %>%
    mutate(predicted_black_stops = predict(fixest_cluster_nb_full,
                                           newdata = new_data_prop) * black,
           type = factor(type, levels = c("min", "p25","med", "p75", "max")),
           variable = "Property Crime Per 10,000")

predict_results <-
    dplyr::select(predict_results_ratio, -nr_officer_black) %>%
    rbind(predict_results_exp, predict_results_prop)

predict_results_ratio %>%
    ggplot(aes(x = type, y = predicted_black_stops)) +
    geom_point(aes(color = year)) +
    geom_line(aes(color = year, group = year)) +
    facet_wrap(~unit, scales = "free") +
    theme_bw() +
    labs(x = "Different Values of Racial Congruence", 
         y = "Predicted Number of Stops of Black Civilians") +
    theme(text = element_text(size = 17))
```

Here is a table trying to link together changes in racial congruence, the predicted number of stops of Black civilians, and changes in the absolute number of Black officers.

```{r}
predict_results_diff <-
    predict_results_ratio %>%
    dplyr::select(unit, year, type, black_ratio, nr_officer_black,
                  predicted_black_stops) %>%
    filter(type == "min" | type == "max") %>%
    pivot_wider(id_cols = c("unit", "year"),
                names_from = type,
                values_from = c("predicted_black_stops", "black_ratio",
                                "nr_officer_black")) %>%
    mutate(prcnt_decrease_stops = (predicted_black_stops_min - predicted_black_stops_max) / predicted_black_stops_min * 100,
           decrease_stops = predicted_black_stops_min - predicted_black_stops_max,
           prcnt_increase_ratio = (black_ratio_max - black_ratio_min) / black_ratio_min * 100,
           prcnt_chg_officer = (nr_officer_black_max - nr_officer_black_min) / nr_officer_black_min * 100,
           chg_officer = nr_officer_black_max - nr_officer_black_min)

predict_table <-
    predict_results_diff %>%
    filter(year == 2014) %>%
    dplyr::select(-prcnt_decrease_stops, -prcnt_increase_ratio,
                  -prcnt_chg_officer, -year) %>%
    mutate(across(matches("stops"), ~round(.x)),
           across(matches("ratio"), ~round(.x, 2))) %>%
    unite("predicted_stops",
          predicted_black_stops_min:predicted_black_stops_max,
          sep = " to ") %>%
    unite("black_ratio",
          black_ratio_min:black_ratio_max,
          sep = " to ") %>%
    unite("nr_officer_black",
          nr_officer_black_min:nr_officer_black_max,
          sep = " to ")
    
kable(predict_table,
      booktabs = T,
      col.names = c("Unit",
                    "Change - Pred. Black Stops",
                    "Min to Max, Racial Con.",
                    "Change - Black Officers",
                    "Diff - Stops",
                    "Diff - Officers"))  %>%
    kable_styling(full_width = F)
```

## Comparing Experience to Racial Congruence

```{r, fig.width = 14, fig.height = 9}
predict_results %>%
    filter(year == 2014) %>%
    ggplot(aes(x = type, y = predicted_black_stops)) +
    geom_point(aes(color = variable)) +
    geom_line(aes(color = variable, group = variable)) +
    facet_wrap(~unit, scales = "free_y") +
    theme_bw() +
    labs(x = "Percentiles", y = "Predicted Number of Stops of Black Civilians") +
    theme(text = element_text(size = 17))
```

Demonstrates how racial congruence compares to the mean amount of experience in a police unit within a given year and how strongly associated each variable is with the decrease in the number of stops of Black civilians. Experience has the stronger association.

## Comparing Percentage of Officers In Unit Who Are Black to Racial Congruence

* Percentage of officers who are Black **and** experience are marginally significant. Percentage of officers who are Black is significant when experience is not included.
* The coefficient is much larger for the percentage of officers who are Black, but that is due to scaling. It has values ranging from 0.02 to 0.69. Racial congruence, on the other hand, ranges from 0.17 to 3.97.
* Racial congruence and work experience in the unit are not strongly correlated (r = `r cor(final_vars$black_ratio, final_vars$mean_years_worked_unit)`).
* Meanwhile, work experience in the unit and the percentage of the police force that is Black are moderately negatively correlated (r = `r cor(final_vars$prcnt_officer_black, final_vars$mean_years_worked_unit)`).

### Using Predicted Number of Black Stops To Contextualize Percentage of Officers Who Are Black vs. Racial Congruence

* For almost all units, the association goes in the same direction and the predicted number of Black stops are similar for percentage vs. racial congruence. However, an increasing percentage of the force that is Black seems to be associated with greater reductions in the number of stops of Black civilians as compared to racial congruence.

```{r, fig.width = 14, fig.height = 9}
new_data_prcnt <-
    final_vars %>%
    group_by(unit, year, black) %>%
    summarise(prcnt_officer_FEMALE = mean(prcnt_officer_FEMALE),
              mean_years_worked_unit = mean(mean_years_worked_unit),
              violent_cr_capita = mean(violent_cr_capita),
              property_cr_capita = mean(property_cr_capita),
              log_total_officers = mean(log_total_officers),
              black = mean(black),
              min = min(prcnt_officer_black),
              max = max(prcnt_officer_black),
              med = median(prcnt_officer_black),
              p25 = quantile(prcnt_officer_black)[["25%"]],
              p75  = quantile(prcnt_officer_black)[["75%"]]) %>%
    pivot_longer(c("min", "max", "med", "p25", "p75"),
                 names_to = "type",
                 values_to = "prcnt_officer_black") %>%
    ungroup() %>%
    full_join(nr_officers, by = c("unit", "year", "type"))

predict_results_prcnt <-
    new_data_prcnt %>%
    mutate(predicted_black_stops = predict(fixest_cluster_nb_prcnt_full,
                                           newdata = new_data_prcnt) * black,
           type = factor(type, levels = c("min", "p25", "med", "p75", "max")),
           variable = "Percentage of Officers Who Are Black") %>%
    select(unit, year, type, variable, predicted_black_stops)

predict_results_compare <-
    predict_results %>%
    filter(variable == "Black Racial Congruence") %>%
    select(unit, year, type, variable, predicted_black_stops) %>%
    rbind(predict_results_prcnt)

predict_results_compare %>%
    filter(year == 2014) %>%
    ggplot(aes(x = type, y = predicted_black_stops)) +
    geom_point(aes(color = variable)) +
    geom_line(aes(color = variable, group = variable)) +
    facet_wrap(~unit, scales = "free_y") +
    theme_bw() +
    labs(x = "Percentiles", y = "Predicted Number of Stops of Black Civilians") +
    theme(text = element_text(size = 16))
```

## Dividing up the sample by Percentage of the Population That Is White

* First, divide up all units into 3 groups. High, medium, and low percentages of the civilian population that is White.
* Second, graph the relationship between racial congruence and the stopping rate of Black civilians for each of the 3 groups:
    * For those units with a high and medium percentage of the civilian population that is White, increasing racial congruence is associated with a higher stopping rate between units.
    * For those units with a low percentage of the civilian population that is White, increasing racial congruence is associated with a lower stopping rate between units.

```{r, fig.width = 12, fig.height = 7}
final_vars_list <-
    final_vars %>%
    mutate(white_percentile = as.numeric(ntile(prcnt_civ_white, 3)),
           white_percentile = case_when(unit == 17 ~ 2,
                                       unit == 4 ~ 1,
                                       T ~ white_percentile)) %>%
    group_split(white_percentile)

map(final_vars_list, function(df) {
    
    ggplot(df, aes(y = black_stop_rate, x = prcnt_officer_black)) +
    geom_point(aes(color = unit)) +
    geom_smooth(method = MASS::glm.nb,
                formula = y ~ x,
                se = F,
                aes(group = unit, color = unit)) +
    geom_smooth(method = MASS::glm.nb,
                formula = y ~ x,
                se = F,
                aes(group = 1)) +
    theme_bw() +
    labs(x = "Black Racial Congruence",
         y = "Stops of Black Civilians Per 10,000") +
    theme(text = element_text(size = 14))
})
```

## Run Regression Analyses on Each Group

* Suprisingly for units with a low percentage of the population that is White, there appears to be no relationship between racial congruence and the stopping rate of Black civilians within unit. The standard error is so high relative to the estimate (which is also suprisingly very high) that there does not appear to be a relationship.
* In fact, there is no statistically significant relationship between racial congruence and the stopping rate of Black civilians in any of the groups. For areas with a medium percentage of the population that is White, there a negative, marginally significant relationship.
* Experience is negative and significant, though, in areas with low and high percentages of the population that is White. The relationship is negative but not significant in areas with a medium percentage of the population that is White.

```{r}
regression_groups <-
    map(final_vars_list, function(df) {
        
        fixest_cluster_nb_full <-
            femlm(black_stops ~ black_ratio + mean_years_worked_unit +
              violent_cr_capita + property_cr_capita + log_total_officers +
              year + offset(log(black)) | unit,
          family = "negbin",
          data = df)
})

names(regression_groups) <- rep("Stops of Black Civilians", 3)

offset_row <- tibble(term = c("", "Offset - Black Pop."),
                     `Model 1` = c("Low White Pop.", "Yes"),
                     `Model 2` = c("Medium White Pop.", "Yes"),
                     `Model 3` = c("High White Pop.", "Yes"))
attr(offset_row, "position") <- c(1, 10)

modelsummary(regression_groups,
             coef_omit = "(Intercept)|theta",
             coef_rename = c(black_ratio = "Black Racial Congruence",
                             mean_years_worked_unit = "Years Worked In Unit (Mean)",
                             prcnt_officer_black = "Percentage of Officers Who Are Black",
                             violent_cr_capita = "Violent Crime Per 10,000",
                             property_cr_capita = "Property Crime Per 10,000",
                             log_total_officers = "Log of the Total Number of Officers"),
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             stars = T,
             gof_omit = "R2$|R2 Adj|R2 Within",
             exponentiate = T,
             output = "latex",
             add_rows = offset_row,
             notes = c("Standard Errors in parentheses. Coefficients are incident rate ratios.",
                       "P-values are denoted by symbols: + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"))
```
