---
title: "Arrests Codebook"
format: pdf
---

```{r, include = F}
library(here)
library(readr)
library(dplyr)
library(purrr)
library(knitr)
library(flextable)
opts_chunk$set(message = F, echo = F, warning = F)
source(here("create_codebooks", "src", "generate_descriptive_tables.R"))
```

```{r}
arrests <- 
    read_csv(
        here("join_outcomes_assignments", "input", "arrests.csv.gz"),
        col_types = 
            cols_only(
                date = col_date(format = "%Y-%m-%d"),
                hour = "n",
                crime_code = "c",
                statute_description = "c",
                lat = "n",
                lon = "n",
                district = "c",
                civ.race = "c",
                civ.gender = "c",
                civ.age = "n",
                arrest_id = "c",
                officer_id = "c",
                month = col_date(format = "%Y-%m-%d"),
                civilian_race_short = "c"
            )
    )
```

## Arrests

* **date**: Date of the arrest in YYYY-MM-DD format.
* **hour**: Hour of the day when the arrest took place rounded to the nearest hour in military time.
* **crime_code**: The suspected crime type causing the arrest.
* **statute_description**: More detailed categories describing what specific statute was suspected to have been violated.
* **lat**: Latitude of the arrest.
* **lon**: Longitude of the arrest.
* **district**: Police district where the arrest took place.
* **civ.race**: Race of the civilian.
* **civ.gender**: Sex of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **arrest_id**: Identifier for each arrest.
* **officer_id**: Unique identifier for each officer.
* **month**: Month of the arrest in YYYY-MM-DD format. The day is always the first day of the month.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* The unit of analysis is a unique officer involved in an arrest. Each row can be uniquely identified by **officer_id** and **arrest_id**.
* Each arrest involves only one civilian, but they can involve multiple officers. It is possible multiple arrests are all a part of one larger incident involving multiple civilians. This can be investigated by examining arrests that took place in the same location at the same time involving the same officers.
* **Number of rows**: `r nrow(arrests)`
* **Number of unique arrests**: `r length(unique(arrests$arrest_id))`
# **Number of unique officers**: `r length(unique(arrests$officer_id))`

> "Arrests for municipal code violations and outstanding warrants are categorized as 'other'". page 8 of Appendix Section S1.5 in Ba et al. 2021.

```{r}
descriptive_table_c_l <-
    arrests %>%
    select((where(is.character) | where(is.logical)) & !matches("id|statute")) %>%
    describe_columns_c_l(nmax = Inf)

descriptive_table_statute <-
    arrests %>%
    select(matches("statute")) %>%
    describe_columns_c_l(nmax = 10)

flextable(bind_rows(descriptive_table_c_l, descriptive_table_statute)) %>%
    align(align = "left", part = "all" ) %>%
    set_table_properties(layout = "autofit") %>%
    fontsize(size = 10, part = "all")
```

```{r}
descriptive_table_n <-
    arrests %>%
    select(where(is.numeric)) %>%
    describe_columns_n()

flextable(descriptive_table_n) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 3)) %>%
    fontsize(size = 8, part = "all")
```

```{r}
descriptive_table_id <-
    arrests %>%
    select(matches("id")) %>%
    describe_columns_id()

flextable(descriptive_table_id)
```

```{r}
descriptive_table_date_num <-
    arrests %>%
    select(matches("date")) %>%
    describe_columns_date_num()

flextable(descriptive_table_date_num) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 8)) %>%
    fontsize(size = 9, part = "all")
```

```{r}
descriptive_table_date_category <-
    arrests %>%
    select(matches("date")) %>%
    describe_columns_date_category()

descriptive_table_hour <-
    arrests %>%
    select(hour) %>%
    describe_columns_c_l(nmax = Inf) %>%
    select(-nr_unique) %>%
    rename(n = nr) %>%
    mutate(
        unit_of_time = "hour",
        value = as.numeric(value)
    )

flextable(bind_rows(descriptive_table_date_category, descriptive_table_hour)) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 10))
```
