---
title: "Use of Force Codebook"
format: pdf
---

```{r, include = F}
library(here)
library(readr)
library(dplyr)
library(purrr)
library(knitr)
library(flextable)
opts_chunk$set(message = F, echo = F, warning = F)
source(here("create_codebooks", "src", "generate_descriptive_tables.R"))
```

```{r}
force <-
    read_csv(
        here("join_outcomes_assignments", "input", "force.csv.gz"),
        col_types = 
            cols_only(
                date = col_date(format = "%Y-%m-%d"),
                time = col_datetime(),
                district = "c",
                lat = "n",
                lon = "n",
                civ.race = "c",
                civ.gender = "c",
                civ.age = "n",
                civ.injured = "c",
                force_id = "c",
                officer_id = "c",
                month = col_date(format = "%Y-%m-%d"),
                civilian_race_short = "c",
                hour = "n"
            )
    )
```

## Uses of Force

* **date**: Date of the use of force in YYYY-MM-DD format.
* **time**: Time of the use of force in YYYY-MM-DD HH:MM:SS format.
* **district**: Police district where the use of force took place.
* **lat**: Latitude of the use of force.
* **lon**: Longitude of the force.
* **civ.race**: Race of the civilian.
* **civ.gender**: Sex of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **civ.injured**: Was the civilian injured?
* **force_id**: Unique identifier for each use of force incident.
* **officer_id**: Unique identifier for each officer.
* **month**: Month of the arrest in YYYY-MM-DD format. The day is always the first day of the month.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* **hour**: Hour of the day when the use of force took place rounded to the nearest hour in military time.
* The unit of analysis is a use of force incident. Only one police officer is listed for each use of force incident. Each row can be uniquely identified by **force_id**.
* **Number of uses of force**: `r nrow(force)`
* **Number of officers**: `r length(unique(force$officer_id))`

```{r}
descriptive_table_c_l <-
    force %>%
    select((where(is.character) | where(is.logical)) & !matches("id")) %>%
    describe_columns_c_l(nmax = Inf)

flextable(descriptive_table_c_l) %>%
    align(align = "left", part = "all" ) %>%
    set_table_properties(layout = "autofit") %>%
    fontsize(size = 10, part = "all")
```

```{r}
descriptive_table_n <-
    force %>%
    select(where(is.numeric)) %>%
    describe_columns_n()

flextable(descriptive_table_n) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 3)) %>%
    fontsize(size = 8, part = "all")
```

```{r}
descriptive_table_id <-
    force %>%
    select(matches("id")) %>%
    describe_columns_id()

flextable(descriptive_table_id)
```

```{r}
descriptive_table_date_num <-
    force %>%
    select(matches("time")) %>%
    describe_columns_date_num()

flextable(descriptive_table_date_num) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 3)) %>%
    fontsize(size = 8, part = "all")
```

```{r}
descriptive_table_date_category <-
    force %>%
    select(matches("time")) %>%
    describe_columns_datetime_category()

flextable(descriptive_table_date_category) %>%
    align(align = "left", part = "all") %>%
    set_table_properties(layout = "autofit", opts_pdf = list(tabcolsep = 10))
```
