---
title: "Shift Assignment Codebook"
format: pdf
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))
```

```{r}
assignments <-
    my_read_csv(here("create-officer-assignments", "input", "assignments.csv"))

# Summarize data
summary <- GetSummary(assignments)
```

## Shift Assignments

* **officer_id**: Unique identifier for each officer.
* **month**: Month of the shift in YYYY-MM-DD format. The day is always the first day of the month.
* **rank**: Rank of the officer assigned to the shift.
* **unit**: Unit of the officer assigned to the shift.
* **date**: Date of the shift in YYYY-MM-DD format.
* **shift**: The shift the officer is assigned to.
* **start_time**: Hour start time of the shift in military time.
* **end_time**: Hour end time of the shift in military time.
* **weekday**: Day of the week of the shift.
* **beat_assigned**: The beat the officer is assigned to.
* **appointed_month**: The month and year the officer was made an officer.
* **months_from_start**: The number of months between the officer's appointment date and their shift date.
* **months_from_start_sq**: The number of months between the officer's appointment date and their shift date, squared.
* **duration**: Length of the shift in hours.
* Uniquely identified by **officer_id** and **date**. The unit of observation is a specific shift for a specific officer.
* **Number of shift assignments**: `r nrow(assignments)`

\newpage

```{r, results = "asis"}
summary[["numeric"]] %>%
    kable(format = "latex") %>%
    kable_styling(latex_options = "scale_down")
```

\newpage

```{r, results = "asis"}
summary[["other"]] %>% kable(format = "latex")
```

\newpage

```{r, results = "asis"}
summary[["factor"]] %>% kable(format = "latex", longtable = T)
```
