---
title: "Officer Codebook"
format: pdf
editor: visual
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))
```

```{r}
officers <-
    my_read_csv(here("create-officer-assignments", "input", "officers.csv"))

summary <- GetSummary(officers)
```

## Officer Roster

-   **birth_year**: Birth year of the officer.
-   **appointed_month**: The month and year the officer was made an officer in YYYY-MM-DD format. The day is always the first day of the month.
-   **officer_id**: Unique identifier for each officer.
-   **officer_race**: Race of the officer.
-   **officer_gender**: Sex of the officer.
-   **spanish**: Does the officer speak Spanish or not?
-   Uniquely identified by **officer_id**. The unit of observation is an officer.
-   **Number of officers**: `r nrow(officers)`

### Creating the officer roster

> "The administrative data from the CPD used in this study span multiple datasets collected in collaboration with the Invisible Institute, Sam Stecklow, and Emma Herman over the course of three years (2016-2019). We obtained these records from the Chicago Police Department or Chicago Department of Human Resources via Freedom of Information Act (FOIA) or through court ordered releases stemming from requests made by Invisible Institute and Jaime Kalven. CPD provided the following data: rosters of all available current and past officers up to 2018, unit history data for individual officers from the 1930s to 2016, Tactical Response Reports from 2004 to 2018 (i.e. use of force reports), and arrest data with arresting officers and arrestee demographic information from 2001 to 2017. The Chicago Department of Human Resources provided data on officers' language skills up to 2019. We supplement our core data with data on 'Stop, Question and Frisk' (SQF) activity between 2012-2015, which was shared by the Lucy Parson's Lab. Finally, the Automated Daily Attendance and Assignment sheet data for each police district between 2012 and 2015 was obtained via a FOIA request to the CPD and shared by Rachel Ryley." pages 5-6 of Appendix Section S1.2 in Ba et al. 2021.

> "These data and others have been used to construct rich profiles of Chicago Police Officers. While no file contains a unique identifier (star numbers change over time, names are common, etc.), we constructed unique officer profiles through a successive merge process described here. Each file contains some identifying information such as of demographic data (birth year, race, gender) or other characteristics (name, start/badge number, appointed date, resignation date, current unit). We used these identifying characteristics to first de-duplicate officers within a file and to then merge to pre-existing officer data with inter-file unique identifiers. The merging process itself is an iterative-pairwise matching method, where the officers in each dataset are repeatedly merged on identifying characteristics and any successful 1-to-1 match in a round removes the matched officers from the next round of merging." page 6 of Appendix Section S1.2 in Ba et al. 2021.

### Officer Race

> "We determine race/ethnicity of CPD officers based on demographic data obtained from the CPD through FOIA. The CPD usually classifies race/ethnicity in at most 7 mutually exclusive groups: White/Caucasian, White Hispanic, Black/African American, Black Hispanic, Asian/Pacific Islander, Native American/Native Alaskan, and unknown/missing. However, there are inconsistencies in how races and ethnicities are coded across files. For example, some files do not include 'Black Hispanic' as a racial category (very few officers are ever classified as Black Hispanic), and some files contain outdated racial categories which we update to the best of our ability. For consistency, we classify 'Hispanic' and 'White Hispanic' as 'Hispanic'; 'Black' and 'Black Hispanic' (rare cases) as 'Black.' 'White' in our analysis refers to non-Hispanic White. If an officer has multiple races associated with them across different datasets, we aggregate by most common non-missing races." page 5 of Appendix Section S1.1 in Ba et al. 2021.

\newpage

```{r, results = "asis"}
map(summary, kable, format = "latex")
```
