---
title: "Arrests Codebook"
format: pdf
---

```{r, include = F}
library(knitr)
opts_chunk$set(message = F, echo = F, warning = F)

library(here)
source(here("functions.R"))
```

```{r}
arrests <- my_read_csv(here("merge-arrests-shifts", "input", "arrests.csv"))
summary <- GetSummary(arrests)
```

## Arrests

* **date**: Date of the arrest in YYYY-MM-DD format.
* **hour**: Hour of the day when the arrest took place rounded to the nearest hour in military time.
* **crime_code**: The suspected crime type causing the arrest.
* **statute_description**: More detailed categories describing what specific statute was suspected to have been violated.
* **lat**: Latitude of the arrest.
* **lon**: Longitude of the arrest.
* **district**: Police district where the arrest took place.
* **civ.race**: Race of the civilian.
* **civ.gender**: Sex of the civilian.
* **civ.age**: Age of the civilian at the time of the stop.
* **arrest_id**: Identifier for each arrest.
* **officer_id**: Unique identifier for each officer.
* **month**: Month of the arrest in YYYY-MM-DD format. The day is always the first day of the month.
* **civilian_race_short**: Collapsed version of **civ.race** (less categories).
* The unit of analysis is a unique officer involved in an arrest. Each row can be uniquely identified by **officer_id** and **arrest_id**.
* Each arrest involves only one civilian, but they can involve multiple officers. It is possible multiple arrests are all a part of one larger incident involving multiple civilians. This can be investigated by examining arrests that took place in the same location at the same time involving the same officers.
* **Number of rows**: `r nrow(arrests)`
* **Number of unique arrests**: `r length(unique(arrests$arrest_id))`

> "Arrests for municipal code violations and outstanding warrants are categorized as 'other'". page 8 of Appendix Section S1.5 in Ba et al. 2021.

\newpage

```{r, results = "asis"}
summary[["numeric"]] %>% kable(format = "latex")
summary[["other"]] %>% kable(format = "latex")
summary[["factor"]] %>% kable(format = "latex", longtable = T)
```
